Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms

### 2017-08-15 paper/paper.tex
Things

\begin{document}
\title[Observations from Parallelising Three Subgraph Algorithms]{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}
\titlenote{This work was supported by the Engineering and Physical Sciences
    Research Council [grant numbers EP/K503058/1, EP/M508056, and EP/P026842/1] and the ANR project
SoLStiCe (ANR-13-BS02-0002-01)}

@@

\begin{abstract}
    We discuss our experiences adapting three recent algorithms for maximum common (connected)
    subgraph problems to exploit multi-core parallelism. These algorithms are heavily irregular and
    very sensitive to search order changes, and so do not easily lend themselves to parallel search.
    Nonetheless, our results show that each algorithm can be parallelised successfully, with the
    multi-core algorithms we create being clearly better than the sequential ones. We then look in
    more detail at the results, and discuss how we should be
    measuring speedups for these kinds of algorithm. Because of the difficulty in measuring an
    average speedup when so-called anomalous (superlinear and sublinear) speedups are common, we
    propose a new measure called \emph{aggregate speedup}.
\end{abstract}

@@

\keywords{Maximum common subgraph, Parallel branch and bound}

@@

\begin{figure}[b]
    \centering
    \begin{tikzpicture}[scale=0.33]%{{{
        \begin{scope}
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [color=uofgsandstone!60] (M4) -- (M6);
            \draw [color=uofgsandstone!60] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!60] (M6) -- (M7);
            \draw [color=uofgsandstone!60] (M6) -- (M8);
            \draw [color=uofgsandstone!60] (M6) -- (M9);
            \draw [very thick] (M7) -- (M9);
            \draw [very thick] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=5cm]
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [color=uofgsandstone!60] (M4) -- (M6);
            \draw [color=uofgsandstone!60] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!60] (M6) -- (M7);
            \draw [very thick] (M7) -- (M8);
            \draw [very thick] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=12cm]
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [color=uofgsandstone!60] (M6) -- (M7);
            \draw [very thick] (M6) -- (M8);
            \draw [color=uofgsandstone!60] (M6) -- (M9);
            \draw [color=uofgsandstone!60] (M7) -- (M9);
            \draw [color=uofgsandstone!60] (M8) -- (M9);
        \end{scope}

        \begin{scope}[xshift=17cm]
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
            \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
            \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [very thick] (M3) -- (M5);
            \draw [very thick] (M4) -- (M6);
            \draw [very thick] (M5) -- (M6);
            \draw [very thick] (M3) -- (M1);
            \draw [very thick] (M6) -- (M7);
            \draw [draw=uofgsandstone!60] (M7) -- (M8);
            \draw [draw=uofgsandstone!60] (M8) -- (M9);
        \end{scope}
\end{tikzpicture}
\caption{A maximum common subgraph of the first two graphs has eight vertices, shaded. A maximum
common \emph{connected} subgraph, shown on the right, has only seven vertices.}
\label{figure:mcsexample}
\end{figure}

A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
graph which maps adjacent vertices to adjacent vertices---that is, it \emph{preserves adjacency}. It
is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
non-adjacency. When working with labelled graphs, a subgraph isomorphism must preserve labels, and
on directed graphs, it must preserve orientation. A \emph{common induced subgraph} of two graphs $G$
and $H$ is a pair of induced subgraph isomorphisms from a pattern graph $P$, one to $G$ and one to
$H$. A \emph{maximum common induced subgraph} is one with as many vertices as possible.

Finding a maximum common subgraph is the key step in measuring the similarity or difference between
two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
the difference between two graphs, we find what they have in common, and then take everything left
over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
represent molecules, but also in applications including computer vision
\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, deanonymising datasets
\citep{o:SharadD13}, the analysis of source code \citep{DBLP:journals/tkde/DjokoCH97} and binary
programs \citep{DBLP:conf/icics/GaoRS08}, in
malware detection \citep{DBLP:journals/compsec/ParkRS13}, social network analysis
\citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
\citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
\citep{DBLP:journals/pr/LuRS91}.  A common variant of the problem requires a largest
\emph{connected} subgraph
\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.
As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
from the maximum common subgraph.

Although both variants are NP-hard, some progress has been made towards solving the problem in
practice.  This paper looks at three recent branch and bound algorithms for maximum common
(connected) induced subgraph problems, each of which is the state of the art for certain classes of
instance. We discuss our experiences in adding parallel tree-search to these three algorithms. In
each case, our results show that the parallel version of the algorithm is clearly better than the
sequential version, although a closer look at the results shows many nuances.

@@

The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and $H$ is an
injective \emph{partial} mapping from $G$ to $H$ which preserves both adjacency and non-adjacency.
Hence we pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other
graph is called the \emph{target}. The model then follows from this new definition: for each vertex
in the pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus
an additional value $\bot$ representing an unmapped vertex. We then have three sets of constraints.
The first set says that for each pair of adjacent vertices in the pattern (that is, for each edge in
the pattern), if neither of these vertices are mapped to $\bot$ then these vertices must be mapped
to an adjacent pair of target vertices. The second set is similar, but looks at non-adjacent pairs
(or non-edges).  Finally, the third set ensures injectivity, by enforcing that the variables must be
all different except when using $\bot$. This final set of constraints may either be implemented
using binary constraints between all pairs of variables, or a special global ``all different except
$\bot$'' propagator \citep{DBLP:conf/cp/PetitRB01}. The objective is simply to find an assignment of
values to variables, maximising the number of variables not set to $\bot$. The state of the art for
this technique is a dedicated implementation of a forward-checking branch and bound search over this
model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}.

@@

A different take on the constraint programming model is presented by
\citet{DBLP:conf/aaai/HoffmannMR17}. They approach the maximum common subgraph via the
subgraph isomorphism problem, asking the question ``if a pattern graph cannot be found in the
target, how much of the pattern graph can be found?''. The $k{\downarrow}$ algorithm tries to solve
the subgraph isomorphism problem first for $k=0$ (asking whether the whole pattern graph can be found in the
target). Should that be not satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot
be matched), and should that also not be satisfiable, it iteratively increases $k$ until the result
is satisfiable. The advantage of this approach is that strong invariants using paths and the degrees
of vertices may be used to prune large portions of the search space.

@@

Most of the benchmark instances we will use come from a standard database for maximum common
subgraph problems \citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}. This benchmark
set can be used in a number of ways, for different variants of the problem. We use it to create five
families of instance, as follows:

@@

The clique and $k{\downarrow}$ algorithms already make use of fine granularity bit-parallelism. To
introduce more coarse-grained thread parallelism, we will parallelise the search process: viewing
backtracking search as forming a tree, we can explore different portions of the tree using different
threads. We use a shared incumbent, so better solutions found by one thread can be used others
immediately. This technique has a long history \citep{o:BaderHC05}. Of particular
interest to us are so-called \emph{anomalies}
\citep{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}:
because we are not performing a fixed amount of work, we should have no expectation of a linear
speedup, and instead we could see a sublinear speedup (much less than $n$ from $n$ processors) or a
superlinear speedup (much more than $n$ from $n$ processors). An absolute slowdown (speedup much
less than 1) is also possible under some models.

@@

Thread-parallel versions of state-of-the-art maximum clique algorithms have already been produced
\citep{DBLP:journals/algorithms/McCreeshP13,DBLP:journals/jcisd/DepolliKRTJ13,DBLP:conf/icde/XiangGA13}.
\citet{DBLP:journals/topc/McCreeshP15} compare several of these approaches, and make an important
observation: although work balance is a problem due to the irregularity of the search tree, often
the interaction between search order and parallel work decomposition is the dominating factor in
determining speedups. They explain why anomalies are in fact common in practice: many clique problem
instances benefit immensely from having found a strong incumbent, but have solutions which are
either unique or rare, and are hard to find. They propose a work splitting mechanism which offsets
anomalies, guaranteeing reproducibility (two runs with the same instance on the same hardware will
give similar runtimes), scalability (increasing the number of cores cannot make things worse), and
no absolute slowdowns.  Additionally, this mechanism explicitly offsets the commitment to early
branching choices, where search ordering heuristics are most likely to be inaccurate
\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}.

We will use this mechanism for our experiments.  The clique-based maximum common subgraph algorithm
effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
described by \citet{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that may be parallelised in
exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
rather than the original three, since maximum common subgraph instances appear to be even more
irregular than normal clique problem instances.

@@

A similar approach may be used for the $k{\downarrow}$ algorithm. Although it is not quite a
conventional branch and bound algorithm, each individual $k$ pass is a tree search, and may be
parallelised. For each pass, we use the same work splitting mechanism as in the clique algorithm,
starting by splitting only at the top level of search to explicitly introduce diversity, and then
iteratively increasing the splitting depth as additional work is needed (up to a limit of five
levels deep).

Because the $k{\downarrow}$ algorithm uses a conventional constraint programming domain store, there
is no need to use recomputation; the state is naturally copied at each branching point.

@@

We perform our experiments on systems with dual Intel Xeon E5-2640 v2 processors and 64GBytes RAM,
running Ubuntu 14.04, with GCC 5.3.0 as the compiler. Each machine has a total of sixteen cores, and
hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted, but
we do not expect speedups of 32 even in the best case.

?? Hundred second timeout.

@@

>>>>

\begin{document}
\title[Observations from Parallelising Three Subgraph Algorithms]{Observations from Parallelising Three Maximum Common (Connected) Subgraph Algorithms}

@@

\begin{abstract}
    We discuss our experiences adapting three recent algorithms for maximum common (connected)
    subgraph problems to exploit multi-core parallelism. These algorithms are heavily irregular and
    very sensitive to search order changes, and so do not easily lend themselves to parallel search.
    Nonetheless, our results show that each algorithm can be parallelised successfully, with the
    multi-core algorithms we create being clearly better than the sequential ones. We then look in
    more detail at the results, and discuss how we should be
    measuring speedups for this kind of algorithm. Because of the difficulty in measuring an
    average speedup when so-called anomalous (superlinear and sublinear) speedups are common, we
    propose a new measure called \emph{aggregate speedup}.
\end{abstract}

@@

\keywords{Maximum common subgraph, Maximum clique, Parallel branch and bound, Constraint
programming, Irregular parallelism, Parallel tree search}

@@

% \begin{figure}[b]
%     \centering
%     \begin{tikzpicture}[scale=0.33]%{{{
%         \begin{scope}
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
% 
%             \draw [very thick] (M1) -- (M2);
%             \draw [very thick] (M2) -- (M4);
%             \draw [very thick] (M3) -- (M5);
%             \draw [color=uofgsandstone!60] (M4) -- (M6);
%             \draw [color=uofgsandstone!60] (M5) -- (M6);
%             \draw [very thick] (M3) -- (M1);
%             \draw [color=uofgsandstone!60] (M6) -- (M7);
%             \draw [color=uofgsandstone!60] (M6) -- (M8);
%             \draw [color=uofgsandstone!60] (M6) -- (M9);
%             \draw [very thick] (M7) -- (M9);
%             \draw [very thick] (M8) -- (M9);
%         \end{scope}
% 
%         \begin{scope}[xshift=5cm]
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
%             \node[draw, circle, fill=magmapurple, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
% 
%             \draw [very thick] (M1) -- (M2);
%             \draw [very thick] (M2) -- (M4);
%             \draw [very thick] (M3) -- (M5);
%             \draw [color=uofgsandstone!60] (M4) -- (M6);
%             \draw [color=uofgsandstone!60] (M5) -- (M6);
%             \draw [very thick] (M3) -- (M1);
%             \draw [color=uofgsandstone!60] (M6) -- (M7);
%             \draw [very thick] (M7) -- (M8);
%             \draw [very thick] (M8) -- (M9);
%         \end{scope}
% 
%         \begin{scope}[xshift=12cm]
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M8) at ($(330:1.5) + (M6)$) {\phantom{0}};
%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(270:1.5) + (M6)$) {\phantom{0}};
% 
%             \draw [very thick] (M1) -- (M2);
%             \draw [very thick] (M2) -- (M4);
%             \draw [very thick] (M3) -- (M5);
%             \draw [very thick] (M4) -- (M6);
%             \draw [very thick] (M5) -- (M6);
%             \draw [very thick] (M3) -- (M1);
%             \draw [color=uofgsandstone!60] (M6) -- (M7);
%             \draw [very thick] (M6) -- (M8);
%             \draw [color=uofgsandstone!60] (M6) -- (M9);
%             \draw [color=uofgsandstone!60] (M7) -- (M9);
%             \draw [color=uofgsandstone!60] (M8) -- (M9);
%         \end{scope}
% 
%         \begin{scope}[xshift=17cm]
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M1) at (90:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M2) at (150:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M3) at (30:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M4) at (210:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M5) at (330:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M6) at (270:1.5) {\phantom{0}};
%             \node[draw, circle, fill=magmaorange, inner sep=0.5pt, font=\normalsize] (M7) at ($(210:1.5) + (M6)$) {\phantom{0}};
%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M8) at ($(270:1.5) + (M6)$) {\phantom{0}};
%             \node[draw, circle, draw=uofgsandstone!60, fill=white, inner sep=0.5pt, font=\normalsize] (M9) at ($(330:1.5) + (M8)$) {\phantom{0}};
% 
%             \draw [very thick] (M1) -- (M2);
%             \draw [very thick] (M2) -- (M4);
%             \draw [very thick] (M3) -- (M5);
%             \draw [very thick] (M4) -- (M6);
%             \draw [very thick] (M5) -- (M6);
%             \draw [very thick] (M3) -- (M1);
%             \draw [very thick] (M6) -- (M7);
%             \draw [draw=uofgsandstone!60] (M7) -- (M8);
%             \draw [draw=uofgsandstone!60] (M8) -- (M9);
%         \end{scope}
% \end{tikzpicture}
% \caption{A maximum common subgraph of the first two graphs has eight vertices, shaded. A maximum
% common \emph{connected} subgraph, shown on the right, has only seven vertices.}
% \label{figure:mcsexample}
% \end{figure}

A \emph{subgraph isomorphism} is an injective mapping from a \emph{pattern} graph to a \emph{target}
graph which maps adjacent vertices to adjacent vertices---that is, it \emph{preserves adjacency}. It
is \emph{induced} if additionally it maps non-adjacent vertices to non-adjacent vertices, preserving
non-adjacency. When working with labelled graphs, a subgraph isomorphism must also preserve labels, and
on directed graphs, it must preserve orientation. A \emph{common induced subgraph} of two graphs $G$
and $H$ is a pair of induced subgraph isomorphisms from a pattern graph $P$, one to $G$ and one to
$H$. A \emph{maximum common induced subgraph} is one with as many vertices as possible. (The
\emph{maximum common partial subgraph} problem is non-induced, with as many edges as
possible; this paper discusses only induced problems.)

Finding a maximum common subgraph is the key step in measuring the similarity or difference between
two graphs \citep{DBLP:journals/prl/Bunke97,DBLP:journals/prl/FernandezV01,o:Kriege15}: to determine
the difference between two graphs, we find what they have in common, and then take everything left
over. Because of this, maximum common subgraph problems frequently arise in biology and chemistry
\citep{DBLP:journals/jcamd/RaymondW02a,o:EhrlichR11,DBLP:journals/dam/GayFMSS14} where graphs
represent molecules or reactions, and also in computer vision
\citep{DBLP:journals/jair/CookH94,DBLP:conf/gbrpr/CombierDS13}, computer-aided manufacturing
\citep{o:LuoWSN17}, crisis management \citep{o:DelavalladeFLL16}, the analysis of programs and
malware \citep{DBLP:conf/icics/GaoRS08,DBLP:journals/compsec/ParkRS13}, social network analysis
\citep{DBLP:journals/tkde/FangYZZ15}, graph database query explanations
\citep{DBLP:journals/jcss/VasilyevaTBL16}, and in character recognition problems
\citep{DBLP:journals/pr/LuRS91}.  A common variant of the problem requires a largest
\emph{connected} subgraph
\citep{DBLP:journals/tcs/Koch01,DBLP:journals/jcamd/RaymondW02a,DBLP:conf/mco/VismaraV08,o:EhrlichR11,o:LuoWSN17}.

% As illustrated in \cref{figure:mcsexample}, a maximum common connected subgraph cannot be determined
% from the maximum common subgraph.

Although both variants are NP-hard, recently progress has been made towards solving the problem in
practice.  This paper looks at three recent branch and bound algorithms for maximum common
(connected) induced subgraph problems, each of which is the state of the art for certain classes of
instance. We discuss our experiences in adding parallel tree-search to these three algorithms. In
each case, our results show that the parallel version of the algorithm is clearly better than the
sequential version, although a closer look at the results shows many nuances.

@@

The maximum common subgraph problem may be reformulated as a constraint optimisation problem, as
follows.  Observe that an equivalent definition of a common subgraph of graphs $G$ and $H$ is an
injective \emph{partial} mapping from $G$ to $H$ which preserves both adjacency and non-adjacency.
Hence we pick whichever input graph has fewer vertices, and call it the \emph{pattern}; the other
graph is called the \emph{target}. The model then follows from this new definition: for each vertex
in the pattern, we create a variable, whose domain ranges over each vertex in the target graph, plus
an additional value $\bot$ representing an unmapped vertex. We then have three sets of constraints.
The first set says that for each pair of adjacent vertices in the pattern (that is, for each edge in
the pattern), if neither of these vertices are mapped to $\bot$ then these vertices must be mapped
to an adjacent pair of target vertices. The second set is similar, but looks at non-adjacent pairs
(or non-edges).  Finally, the third set ensures injectivity, by enforcing that the variables must be
all different except when using $\bot$. This final set of constraints may either be implemented
using binary constraints between all pairs of variables, or a special global ``all different except
$\bot$'' propagator \citep{DBLP:conf/cp/PetitRB01}. The objective is simply to find an assignment of
values to variables, maximising the number of variables not set to $\bot$. The state of the art for
this technique is a dedicated (non-toolkit) implementation of a forward-checking branch and bound
search over this
model \citep{DBLP:conf/cp/NdiayeS11,DBLP:conf/cp/McCreeshNPS16}; attempts to parallelise this
algorithm by static decomposition have had limited success \citep{DBLP:conf/ictai/MinotNS15}.

@@

A different take on the constraint programming model is presented by
\citet{DBLP:conf/aaai/HoffmannMR17}. They approach maximum common subgraph via the
subgraph isomorphism problem, asking the question ``if a pattern graph cannot be found in the
target, how much of the pattern graph can be found?''. The $k{\downarrow}$ algorithm tries to solve
the subgraph isomorphism problem first for $k=0$ (asking whether the whole pattern graph can be found in the
target). Should that be not satisfiable, it tries to solve the problem for $k=1$ (one vertex cannot
be matched), and should that also not be satisfiable, it iteratively increases $k$ until the result
is satisfiable. The advantage of this approach is that strong invariants using paths and the degrees
of vertices may be used to prune large portions of the search space.

@@

Most of the benchmark instances we will use come from a standard database for maximum common
subgraph problems \citep{DBLP:journals/prl/SantoFSV03,DBLP:journals/jgaa/ConteFV07}. This benchmark
set can be used in a number of ways, for different variants of the problem. Following other recent
work \citep{DBLP:conf/cp/McCreeshNPS16,DBLP:conf/aaai/HoffmannMR17,o:McCreeshPT17}, we use it to
create five families of instance, as follows:

@@

The clique and $k{\downarrow}$ algorithms already make use of fine-granularity bit-parallelism. To
introduce more coarse-grained thread parallelism, we will parallelise the search process: viewing
backtracking search as forming a tree, we can explore different portions of the tree using different
threads. We use a shared incumbent, so better solutions found by one thread can be used others
immediately. This technique has a long history \citep{o:BaderHC05}. Of particular
interest to us are so-called \emph{anomalies}
\citep{DBLP:journals/cacm/LaiS84,DBLP:journals/tc/LiW86,DBLP:conf/irregular/BruinKT95}:
because we are not performing a fixed amount of work, we should have no expectation of a linear
speedup, and instead we could see a sublinear speedup (much less than $n$ from $n$ processors, if
speculative work turns out to be wasted) or a superlinear speedup (much more than $n$ from $n$
processors, if a strong incumbent is found more quickly). An absolute slowdown (a speedup much less
than 1) is also possible under some models.

@@

Thread-parallel versions of state-of-the-art maximum clique algorithms have already been produced.
\citet{DBLP:journals/topc/McCreeshP15} compare several of these approaches, and make an important
observation: although work balance is a problem due to the irregularity of the search tree, often
the interaction between search order and parallel work decomposition is the dominating factor in
determining speedups. They explain why anomalies are in fact common in practice: many clique problem
instances benefit immensely from having found a strong incumbent, but have solutions which are
either unique or rare, and are hard to find. They propose a work splitting mechanism which offsets
anomalies, guaranteeing reproducibility (two runs with the same instance on the same hardware will
give similar runtimes), scalability (increasing the number of cores cannot make things worse), and
no absolute slowdowns.  Additionally, this mechanism explicitly offsets the commitment to early
branching choices, where search ordering heuristics are most likely to be inaccurate
\citep{DBLP:conf/ijcai/HarveyG95,DBLP:conf/cp/ChuSS09}. Under this approach, superlinear speedups
are common.

We will use this mechanism for our experiments.  The clique-based maximum common subgraph algorithm
effectively differs only in the preprocessing stage, and the clique-inspired connected algorithm
described by \citet{DBLP:conf/cp/McCreeshNPS16} is sufficiently similar that may be parallelised in
exactly the same way. Based upon preliminary experiments, we use a splitting depth limit of five
rather than the original three, since maximum common subgraph instances appear to give even more
irregular search trees than normal clique problem instances.

@@

A similar approach may be used for the $k{\downarrow}$ algorithm. Although it is not quite a
conventional branch and bound algorithm, each individual $k$ pass is a tree search, and may be
parallelised. For each pass, we use the same work splitting mechanism as in the clique algorithm,
starting by splitting only at the top level of search to explicitly introduce diversity, and then
iteratively increasing the splitting depth as additional work is needed (up to a limit of five
levels deep).  Because the $k{\downarrow}$ algorithm uses a conventional constraint programming
domain store, there is no need to use recomputation; the state is naturally copied at each branching
point.

@@

We perform our experiments on systems with dual Intel Xeon E5-2640 v2 processors and 64GBytes RAM,
running Ubuntu 14.04, with GCC 5.3.0 as the compiler. Each machine has a total of sixteen cores, and
hyper-threading is enabled, so we run experiments with 32 threads except where otherwise noted, but
we do not expect speedups of 32 even in the best case.  We run all our experiments with a one
hundred second timeout for each instance. All of our sequential runtimes are from optimised
implementations which were not designed with parallelism in mind---that is, they are genuine
improvements over the state of the art.

@@

    \vspace*{0.2em}
